\subsubsection{Single-nuclear transcriptomic profiling of human PFC from ABCA7 LoF-variant carriers} 
To investigate the cell type-specific impact of ABCA7 LoF variants in the human brain, we queried whole genome sequences of >1000 subjects from the Religious Order Study or the Rush Memory and Aging Project (collectively known as ROSMAP) for donors with Alzheimer’s disease diagnoses who are carriers of rare damaging variants in ABCA7 that result in a PTC. We identified 12 heterozygous carriers of ABCA7 LoF variants, including splice region variants (c.4416+2T>G and c.5570+5G>C), frameshift variants (p.Leu1403fs and p.Glu709fs), and nonsense ‘stop gained’ variants (p.Trp1245* and p.Trp1085*) (Figure~\ref{fig:main_atlas}A; Figure~\ref{fig:snRNA_cohort}A,B; Data~\ref{data:ptc_variants}). These variants have previously been associated with increased AD risk in genetic association studies (Table~\ref{tab:annotation_abca7})\cite{Steinberg2015-mu,Holstege2022-vp} and are presumed to induce risk via ABCA7 haploinsufficiency\cite{Duchateau2024-rf}. Analysis of published proteomic data\cite{Johnson2020-ip} confirmed that ABCA7 PTC-variant carriers indeed had lower ABCA7 protein levels in the human postmortem PFC compared to non-carriers (p=0.018; N = 180 Controls, 5 ABCA7 LoF; Figure~\ref{fig:main_atlas}B; Figure~\ref{fig:snRNA_cohort}D). 

We next selected 24 ABCA7 PTC non-carrier controls from the ROSMAP cohort that were matched to the ABCA7 LoF variant-carriers based on several potentially confounding variables, including Alzheimer's disease (AD) pathology, age at death, post-mortem intervals, sex, APOE genotype, and cognitive status (Figure~\ref{fig:main_atlas}C,D; Figure~\ref{fig:snRNA_cohort}B,C; Figure~\ref{fig:snRNA_quality_annotation}D,E; Data~\ref{data:cohort_metadata}; Supplementary Text). We confirmed that none of the 36 selected subjects carried damaging variants in other known AD risk genes (TREM2, SORL1, ATP8B4, ABCA1, and ADAM10)\cite{Holstege2022-vp}, and verified ABCA7 genotypes in a subset of ABCA7 LoF carriers and matched controls using Sanger sequencing (Figure~\ref{fig:snRNA_cohort}A).

\newcommand{\quoteE}{\textcolor{blue}{For a subset of the selected samples raw data (fastq files) for snRNAseq of PFC BA10 could be obtained from a previous study (10 non-carrier controls from \cite{Mathys2019-wb}). For the remaining samples, fresh-frozen tissue samples from the BA10 region of the prefrontal cortex (PFC) were obtained for analysis. Single-nucleus RNA sequencing (snRNAseq) was performed using the 10x Genomics Chromium platform. Accurate genotype assignments were confirmed by matching each single-cell library to its corresponding whole genome sequencing data (Figure~\ref{fig:snRNA_cohort}B). Following extensive quality control measures—including detailed analysis and correction of batch effects (Figure~\ref{fig:snRNA_batch_quality}, Figure~\ref{fig:snRNA_quality_annotation}C-E; Methods, Supplementary Text)—our final dataset consisted of 102,710 high-quality cells out of an initial total of 150,456 cells (Figure~\ref{fig:main_atlas}E).\label{quoteE-label}}}
\quoteE This dataset encompassed diverse populations of inhibitory neurons (In, SYT1 & GAD1+), excitatory neurons (Ex, SYT1 & NRGN+), astrocytes (Ast, AQP4+), microglia (Mic, CSF1R+), oligodendrocytes (Oli, MBP & PLP1+), and oligodendrocyte precursor cells (OPCs, VCAN+) (Figure~\ref{fig:main_atlas}E; Figure~\ref{fig:snRNA_quality_annotation}F-I). A small putative vascular cell cluster did not meet our quality thresholds and was excluded from further analysis. Post-quality control, gene expression profiles showed high consistency within cell types (mean correlation 0.95) and were robustly represented across subjects (Figure\ref{fig:snRNA_quality_annotation}J-M).

\subsubsection{Cell type-specific perturbations in the presence of ABCA7 LoF}
To investigate gene expression changes related to ABCA7 loss-of-function (LoF) across major cell types, we identified genes significantly perturbed (p<0.05, linear model; total genes = 2,389) in at least one of six major cell types (Ex, In, Ast, Mic, Oli, or OPC). We controlled for known and unknown covariates and considered only genes detected in >10\% of cells within each specific cell type (see Methods; Data~\ref{data:degs}). Next, \newcommand{\quoteF}{\textcolor{blue}{we visualized these perturbed genes by projecting their high-dimensional perturbation scores (S = sign(log(FC)) × -log10(p-value) for each cell type) onto two dimensions, as shown in Figure~\ref{fig:main_atlas}F. Genes exhibiting similar perturbation patterns across cell types are positioned closer together in this two-dimensional visualization. \label{quoteF-label}}} 
\quoteF

\newcommand{\quoteG}{\textcolor{blue}{The two-dimensional visualization effectively captured the transcriptional landscape of ABCA7 LoF gene changes across all major cell types (Fig1F; Figure~\ref{fig:snRNAseq_gene_scores}A). To summarize this landscape in terms of biological pathways, we grouped genes into clusters based on their positions in the 2D projection and analyzed each cluster for enrichment in biological pathways using the Gene Ontology Biological Process database (Figure~\ref{fig:main_atlas}G; see Methods). This analysis identified several biological pathways broadly affected by ABCA7 LoF in the postmortem human PFC, including pathways related to heat shock response, inflammation, cell dynamics and adhesion, synaptic function, DNA repair, and cellular metabolism (Figure~\ref{fig:main_atlas}G; Data~\ref{data:pathway_enrichments}). \label{quoteG-label}}} 
\quoteG

Decomposing the ABCA7 LoF transcriptional signature revealed both common and unique cell type-specific gene perturbation patterns across major cell types in the PFC (Figure 1G; Figure~\ref{fig:snRNAseq_gene_scores}D; Data~\ref{data:degs}-5; Methods). For instance, microglia showed a down-regulated heat shock response (cluster 4), also detected to a lesser extent in neurons and OPCs (FDR-adjusted permutation p-value<0.01, |score|>0.25; Figure~\ref{fig:main_atlas}H). Microglia and astrocytes exhibited up-regulation of inflammatory clusters (cluster 5 and cluster 10, respectively; Figure~\ref{fig:main_atlas}H). OPCs showed perturbations in clusters related to cell motility and endocytosis (clusters 1 and 0; Figure~\ref{fig:main_atlas}H). Both neuronal subtypes demonstrated increased expression of DNA repair-related cluster 9 genes (e.g., NUP205, POLD3, VCP) and decreased expression of clusters 6 and 14, which include cell adhesion and synaptic genes (e.g., SYT11, SHANK3, IFT56, GABRA3, NLGN1; Figure~\ref{fig:main_atlas}H). Additionally, excitatory neurons displayed unique perturbations, specifically up-regulation of mitochondrial and antioxidant genes in cluster 8 (e.g., COX7A2, PARK7), and down-regulation of lipid synthesis and transport genes in cluster 2 (e.g., LDLR, APOL2, NR1H3, ACLY, FITM2; Figure~\ref{fig:main_atlas}H). A detailed comparison of overlapping gene changes across cell types is also provided in Figure~\ref{fig:snRNAseq_gene_scores_2}A,B.

Together, these findings indicate that ABCA7 LoF variants may induce widespread, cell type-specific transcriptional changes in the human PFC. This single-cell atlas provides a rich resource for future studies aiming to elucidate the contributions of individual neural cell types to ABCA7 LoF-driven forms of AD risk. This resource will be made available for exploration via the UCSC Single Cell Browser and for further analysis via Synapse (accession ID: syn53461705).

\subsubsection{ABCA7 is expressed most highly in excitatory neurons}
Our snRNAseq data suggest that excitatory neurons expressed the highest levels of ABCA7, compared to other major cell types in the brain (Figure~\ref{fig:abca7_expression}A).  ABCA7 transcripts were detected (count>0) in ~ 30\% of excitatory neurons and ~ 15\% of inhibitory neurons, while the detection rate was considerably lower (<10\%) for microglia and astrocytes and an order of magnitude lower (<3\%) for oligodendrocytes and OPCs (Figure~\ref{fig:abca7_expression}A, B). We validated this expression pattern in an independent published dataset from\cite{Welch2022-ef} (Table~\ref{tab:external_datasets}), where bulk RNA sequencing of NeuN- (glial) and NeuN+ (neuronal) cell populations derived from six human postmortem temporal cortex samples showed significantly higher ABCA7 levels in the neuronal population versus the glial cell population (p=0.021, paired t-test; Figure~\ref{fig:abca7_expression}C). Several control genes, whose expression patterns in glial versus neuronal cells are well established (ABCA1, APOE, and NEUROD1), had expected expression patterns that matched those in the snRNAseq data (Figure~\ref{fig:abca7_expression}B,C). These results indicate that neurons, particularly excitatory neurons, are the primary ABCA7-expressing cell type in the aged human PFC. Given the relatively higher expression of ABCA7 in excitatory neurons and the evidence of transcriptional perturbations by ABCA7 LoF in this cell type, we focused our subsequent analysis specifically on excitatory neurons.

\subsubsection{ABCA7 LoF perturbations in excitatory neurons}
As an alternative approach to the unsupervised clustering of gene perturbation scores among all cell types, we next used prior knowledge of biological pathway structure to perform an in-depth characterization of perturbed biological processes specifically in ABCA7 LoF excitatory neurons. To this end, we first estimated statistical overrepresentation of biological gene sets (WikiPathways, N pathways = 472) among up and down-regulated genes in ABCA7 LoF excitatory neurons vs controls (by GSEA; Methods). We observed a total of 34 pathways with evidence for transcriptional perturbation at p<0.05 in excitatory neurons (Data~\ref{data:kl_clusters}). Enrichments of these pathways were driven by 268 unique genes (“leading edge” genes\cite{Subramanian2005-pu}; Data~\ref{data:kl_clusters}).   

To extract unique information from leading-edge genes and limit pathway redundancy, we next separated these genes and their associated pathway annotations into non-overlapping groups, formalized as a graph partitioning problem (Figure~\ref{fig:main_neurons}A; Figure~\ref{fig:benchmarking_clustering}; Methods; Supplementary Text). Establishing gene-pathway groupings of approximately equal size revealed eight biologically interpretable “clusters” associated with ABCA7 LoF in excitatory neurons (Figure~\ref{fig:main_neurons}A,B; Data~\ref{data:kl_clusters}). Predominantly, these gene clusters centered around two themes: (1) energy metabolism and homeostasis (C2, C7) and (2) DNA damage (C1, C4, C5), cell stress (C3, C6), and synaptic dysfunction (C0).

Clusters C2 and C7 were primarily defined by genes involved in cellular energetics, including genes related to lipid metabolism, mitochondrial function, and  OXPHOS (Figure~\ref{fig:main_neurons}C,D). Cluster 2, characterized by transcriptional regulators of lipid homeostasis (e.g. NR1H3, ACLY, PPARD), exhibited evidence for down-regulation in ABCA7 LoF and featured pathways related to "SREBP signaling" and "lipid metabolism" (Figure~\ref{fig:main_neurons}B; Data~\ref{data:kl_clusters}). Cluster 7 comprised multiple mitochondrial complex genes (e.g. COX7A2, NDUFV2) responsible for ATP generation from carbohydrate and lipid catabolism and showed up-regulation in ABCA7 LoF (Figure~\ref{fig:main_neurons}B; Data~\ref{data:kl_clusters}).The remaining clusters C0, C1, and C3-C6 were characterized by DNA damage and proteasomal, inflammatory, and apoptotic mediators. Clusters C1, C3, C4, and C5 were up-regulated in ABCA7 LoF excitatory neurons and characterized by pathway terms such as "DNA damage response," "Parkin-Ubiquitin Proteasomal system,” "Signaling pathways in glioblastoma," and "Nucleotide metabolism," respectively (Figure~\ref{fig:main_neurons}B; Data~\ref{data:kl_clusters}). They included up-regulated DNA damage/repair and proteasomal genes (e.g. RECQL, TLK2, BARD1, RBL2, MSH6, PSMD5, PSMD7) and were accompanied by increased expression of inflammatory mediators (e.g. C2, C1S, CLU, PIK3CD in C0, C4, C6) and down-regulated gene signatures related to synaptic function (C0), cytoskeletal function (C3), and apoptosis-related genes (C6) (Figure~\ref{fig:main_neurons}B; Data~\ref{data:kl_clusters}). 

Together, these data suggest that ABCA7 LoF may disrupt energy metabolism in excitatory neurons and that these disruptions coincide with a state of increased cellular stress, characterized by genomic instability and neuronal dysfunction.

\subsubsection{ABCA7 LoF and common missense variants lead to overlapping neuronal perturbations}
\newcommand{\quoteL}{\textcolor{blue}{ABCA7 LoF variants substantially increase AD risk (Odds Ratio = 2.03)\cite{Steinberg2015-mu} but are rare and therefore only contribute to a small portion of AD cases\cite{Duchateau2024-rf}. To evaluate whether ABCA7 LoF transcriptomic effects in neurons generalize to more common, moderate-risk genetic variants in ABCA7, we examined the ROSMAP WGS cohort for carriers of the prevalent ABCA7 missense variant p.Ala1527Gly (rs3752246: Minor Allele Frequency ≈ 0.18; \% carriers ≥1 allele ≈30\%; Figure~\ref{fig:main_excitatory}E). Although Gly1527 is listed as the reference allele, it represents the less common variant associated with increased AD risk (Odds Ratio = 1.15 [1.11-1.18])\cite{Kunkle2019-yo,Holstege2022-vp,Naj2011-bs}. \label{quoteL-label}}} 
\quoteL We identified 133 individuals carrying at least one copy of the p.Ala1527Gly risk variant and 227 non-carriers (Figure~\ref{fig:main_excitatory}F), all with available snRNAseq data from post-mortem PFC\cite{Mathys2023-rs}. We ensured none of these 360 individuals were part of our earlier ABCA7 LoF snRNAseq cohort or carried ABCA7 LoF variants. Using this cohort, we investigated whether excitatory neurons from p.Ala1527Gly carriers exhibited evidence of transcriptomic perturbations in the ABCA7 LoF-associated clusters C0-C7.

Remarkably, all clusters displayed directional trends in p.Ala1527Gly neurons consistent with the mean effect sizes observed in ABCA7 LoF neurons (Figure~\ref{fig:main_excitatory}C,G), while controlling for pathology, age, sex, and other covariates (Methods). Notably, 4 out of 8 clusters exhibited substantial evidence of perturbation in p.Ala1527Gly variant carriers, with perturbation directions aligning with predictions for ABCA7 LoF (Figure~\ref{fig:main_excitatory}G,H). Specifically, we observed an up-regulation in the DNA damage cluster C1 and the proteasomal cluster C3 in p.Ala1527Gly carriers compared to controls, suggesting a similar cell stress and genomic instability signature to ABCA7 LoF carriers (Figure~\ref{fig:main_excitatory}G,H), and a borderline significant up-regulation of the mitochondrial cluster C7, again consistent with ABCA7 LoF (Figure~\ref{fig:main_excitatory}G,H). Intriguingly, the most significantly perturbed cluster was the lipid cluster C2, which exhibited down-regulation (Figure~\ref{fig:main_excitatory}G,H) similar to our observations in ABCA7 LoF carriers. 

\newcommand{\quoteP}{\textcolor{blue}{Because missense variants can alter protein dynamics, and glycine substitutions are known to introduce significantly greater local flexibility compared to alanine, we evaluated whether the convergent transcriptional signature could be explained by structural changes in the ABCA7 protein. To specifically investigate the local structural consequences of the p.Ala1527Gly variant, we leveraged recent advances in protein structure simulations along with newly resolved cryo-EM structures of ABCA7 in both its closed (ATP-bound) (Figure~\ref{fig:main_excitatory}I) and open (ATP-unbound) (Figure~\ref{fig:md_simulations}A) conformations \cite{Le_Thi_My2022-dp,Jumper2021-na}. Molecular dynamics simulations were performed on a 239-residue region of ABCA7, embedded within a lipid bilayer, comparing conditions with and without the p.Ala1527Gly substitution over a 300-ns simulation period (Figure~\ref{fig:main_excitatory}J; Figure~\ref{fig:md_simulations}A; Methods; Supplementary Text).\label{quoteL-labeP}}} 
\quoteP These simulations revealed that the AD risk-associated G1527 mutation in the ATP-bound closed form induced local flexibility, evidencing large structural fluctuations over time, while the A1527 variant had a stabilizing effect on the ATP-bound closed conformation with only minor structural fluctuations (Figure~\ref{fig:main_excitatory}K-M). Both variants displayed stable conformational behavior in the unbound-open conformation (Figure~\ref{fig:md_simulations}A-F). Since the closed ATP-bound conformation is thought to mediate lipid extrusion to the plasma membrane and apolipoproteins\cite{Le_Thi_My2022-dp}, our data indicate that the p.Ala1527Gly substitution significantly impacts local secondary structure  stability and conformational dynamics of ABCA7, potentially affecting its lipid extrusion function.  

Together, our molecular dynamics simulations indicate that the p.Ala1527Gly substitution likely increases AD risk by perturbing ABCA7 structure, and therefore function. As such, our findings imply that rare and common ABCA7 LoF variants may mediate AD risk via similar, ABCA7-dependent mechanisms  and that findings from an in-depth study of ABCA7 LoF may extend to larger portions of the at-risk population.

\subsubsection{Deriving human neurons with ABCA7 LoF variants}
To complement the correlative analyses in ABCA7 LoF human tissue, we next used CRISPR-Cas9 genome editing to generate two isogenic iPSC lines, each homozygous for ABCA7 LoF variants, from a parental line without ABCA7 variants (WT). The first LoF variant, ABCA7 p.Glu50fs*3, was generated by a single base-pair insertion in ABCA7 exon 3, resulting in a PTC early in the ABCA7 gene (Figure~\ref{fig:main_mitochondrial}A; Figure~\ref{fig:differentiating_iPSC_neurons}A-C). This 5'-proximal PTC limits the possibility of producing a truncated protein with partial functionality, generating a near-complete knockout. The second LoF variant, ABCA7 p.Tyr622*, was generated by a single base-pair mutation in ABCA7 exon 15 (Figure~\ref{fig:main_mitochondrial}A; Figure~\ref{fig:lipid_mitochondrial_perturbations}A-C). This PTC re-creates a variant previously observed in patients as associated with AD \cite{Steinberg2015-mu} and is meant to provide clinical context to ABCA7 dysfunction.

We differentiated the isogenic iPSCs into neurons (iNs) through lentiviral delivery of a doxycycline-inducible NGN2 expression cassette as previously described\cite{Ho2016-kz} (Figure~\ref{fig:differentiating_iPSC_neurons}D). At 2 and 4 weeks post-NGN2 induction, cells were immunoreactive for neuronal markers TUJ1 and MAP2 and showed robust neuronal processes by pan-axonal staining (Figure~\ref{fig:differentiating_iPSC_neurons}E). Both WT and ABCA7 LoF lines had the ability to fire action potentials upon current injections (Figure~\ref{fig:iN_phenotypes}A-G). Though the ABCA7 genotype did not affect resting membrane potential (Figure~\ref{fig:iN_phenotypes}E), ABCA7 LoF iNs fired more action potentials and with lesser magnitude of current injection than WT iNs (Figure~\ref{fig:iN_phenotypes}F,G), indicating an ABCA7 LoF hyperexcitability phenotype. 

Previous studies suggest that ABCA7 dysfunction affects amyloid processing\cite{Satoh2015-yu,Sakae2016-uy,Bamji-Mirza2018-xt,Chan2008-qu,De_Roeck2018-zx}. To verify these findings in our ABCA7 LoF iNs, we measured Aβ42 and Aβ40 levels in media from iNs after four weeks of culture. Using enzyme-linked immunosorbent assay (ELISA), we found that ABCA7 LoF iNs exhibited a significant increase in both secreted Aβ42 and Aβ40 compared to wild-type cells (Figure~\ref{fig:differentiating_iPSC_neurons}K). Together, these data (1) confirm successful neuronal differentiation from iPSCs, and (2) show that ABCA7 LoF iNs recapitulate key AD-associated phenotypes of hyperexcitability and amyloid pathology.

\subsubsection{ABCA7 LoF iNs Recapitulate Excitatory Neuronal Transcriptional Signatures}
To investigate whether transcriptional changes associated with ABCA7 loss-of-function (LoF) observed in \emph{postmortem} human neurons are recapitulated in iNs, we performed bulk mRNA sequencing on ABCA7 WT, p.Glu50fs3, and p.Tyr622 iNs (N=2, N=5, and N=5, respectively) after four weeks in culture. Gene perturbation scores (defined as -log10(p-value)*sign(log2(FC))) showed a strong correlation between p.Glu50fs3 vs. WT and p.Tyr622 vs. WT comparisons (Pearson correlation coefficient = 0.84; Figure~\ref{fig:main_mitochondrial}B), indicating consistency in the transcriptional impact of ABCA7 variants.

We next conducted gene set enrichment analysis (GSEA) on the differentially expressed genes from these comparisons, identifying X significantly perturbed pathways in p.Glu50fs3 vs. WT and Y in p.Tyr622 vs. WT (FDR-adjusted p < 0.05; WikiPathways). These pathways were driven by 425 and 335 unique "leading edge" genes, respectively \cite{Subramanian2005-pu}. K/L partitioning of these leading edge genes revealed substantial concordance between the two ABCA7 variant lines, with nearly all clusters (8/9 for p.Tyr622* and 9/10 for p.Glu50fs*3) represented significantly in both comparisons (FDR-adjusted p < 0.05) (Figure~\ref{fig:main_mitochondrial}C; Figure~\ref{fig:bulk_RNAseq_supplement}A,B).

We observed that transcriptional signatures in ABCA7 LoF iNs closely aligned with those identified in postmortem excitatory neurons. Specifically, we found significant overlap in 5 out of 9 Tyr622*-associated clusters (all exhibiting consistent directional trends) and in 9 out of 10 Glu50fs*3-associated clusters, with the majority (7 out of 9) showing concordant directional changes with \emph{postmortem} ABCA7 LoF excitatory neurons from the \emph{postmortem} human brain. (Figure~\ref{fig:main_mitochondrial}D; Figure~\ref{fig:bulk_RNAseq_supplement}C). For clarity and consistency, we renamed the iN clusters to correspond to their matching postmortem clusters whenever significant (FDR<0.05) overlap was identified (Figure~\ref{fig:main_mitochondrial}D,E).

Given the strong similarities between the two LoF lines, we primarily focused our characterization on the patient-variant p.Tyr622*, with complementary analyses for the p.Glu variant provided in supplementary materials (Figure~\ref{fig:bulk_RNAseq_supplement}D). Consistent with findings from postmortem data, p.Tyr622* iNs exhibited downregulated clusters associated with lipid metabolism (PM.0 and T.9) and upregulated clusters related to cell cycle regulation and proteasomal activity (PM.4, T.12, and T.8) compared to WT iNs (Figure~\ref{fig:main_mitochondrial}E). Notably, a mitochondrial cluster (PM.1) demonstrated the most robust overlap with postmortem data and was consistently identified across both the p.Tyr622* and p.Glu50fs lines (Figure~\ref{fig:main_mitochondrial}D; Figure~\ref{fig:bulk_RNAseq_supplement}B,C). Together, these data support a causal relationship between ABCA7 LoF variants and the mitochondrial transcriptional signature observed in postmortem excitatory neurons.

\subsubsection{ABCA7 LoF impairs mitochondrial uncoupling in neurons}
To further characterize mitochondrial alterations in ABCA7 LoF iNs, extending beyond the gene sets used for K/L cluster analysis, we examined the expression of 1,136 mitochondrial genes curated from the MioCarta database in our bulk RNAseq data. Among the most significantly upregulated genes in p.Tyr622* versus WT iNs were components of mitochondrial apoptosis pathways (e.g., CASP3, BID) and oxidative phosphorylation (OXPHOS) subunits (previously identified within cluster PM.X) (Figure~\ref{fig:main_mitochondrial}F; Table~\ref{tab:y622_mito_genes}). Conversely, downregulated genes were significantly enriched (padj < 0.05) for key metabolic processes, including beta-oxidation (ACAD and CPT genes), mitochondrial metabolite transport (SLC25 genes), and oxidative stress detoxification (CAT) (Figure~\ref{fig:main_mitochondrial}F; Table~\ref{tab:y622_mito_genes}). These MioCarta mitochondrial gene expression profiles were highly correlated between p.Tyr622* and p.Glu50fs*3 relative to WT iNs (Figure~\ref{fig:bulk_RNAseq_supplement}E).

To directly assess mitochondrial function in ABCA7 LoF neurons, we measured the oxygen consumption rate (OCR) of WT and ABCA7 LoF iNs over time using the Seahorse metabolic flux assay (Figure~\ref{fig:oxygen_consumption_rates_iPSC_neurons}A,B). The mitochondrial membrane potential (ΔѰm) is maintained by the oxygen-driven movement of protons across the inner mitochondrial membrane during OXPHOS (Figure~\ref{fig:oxygen_consumption_rates_iPSC_neurons}C). \newcommand{\quoteC}{\textcolor{blue}{Measuring OCR in the presence of mitochondrial inhibitors thus provides several functional readouts.Because OCR can be influenced by cell viability and mitochondrial abundance\cite{Divakaruni2014-eq,Gu2021-ms}, we report OCR ratios\cite{Divakaruni2022-rj} for WT, ABCA7 p.Glu50fs30, and ABCA7 p.Tyr622 iNs. First, we assessed spare respiratory capacity by measuring the relative increase in oxygen consumption following pharmacological collapse of the proton gradient\cite{Divakaruni2022-rj} (Figure~\ref{fig:oxygen_consumption_rates_iPSC_neurons}D). We then quantified the proportion of basal oxygen consumption attributed to ATP synthesis versus proton leak (i.e., uncoupled mitochondrial OCR)\cite{Divakaruni2014-eq} (Figure~\ref{fig:main_mitochondrial}E).\label{quoteC-label}}}
\quoteC

While spare respiratory capacity was comparable between WT and ABCA7 LoF iNs (Figure~\ref{fig:oxygen_consumption_rates_iPSC_neurons}F), ABCA7 LoF iNs showed significantly reduced uncoupled mitochondrial respiration (Figure~\ref{fig:main_mitochondrial}G), a finding replicated in two independent experiments (Figure~\ref{fig:oxygen_consumption_rates_iPSC_neurons}G,H). Uncoupled mitochondrial oxygen consumption rates (OCR) in WT iNs (20\%; Figure\ref{fig:main_mitochondrial}G) align with previously reported values for neurons and other cell types \cite{Divakaruni2011-uj,Jekabsons2004-fn}, indicating that ABCA7 LoF iNs exhibit abnormally low mitochondrial uncoupling. Consistent with this observation, expression levels of UCP2 - the primary neuronal mitochondrial uncoupling protein \textcolor{red}{[citation needed]}- were reduced in ABCA7 LoF iNs (Figure~\ref{fig:oxygen_consumption_rates_iPSC_neurons}I) \textcolor{red}{[specify significance]}. 

Because decreased mitochondrial uncoupling often correlates with elevated mitochondrial membrane potential (ΔѰm) \textcolor{red}{[citation needed]}, we next assessed ΔѰm in NeuN-positive soma using the fixable MitoHealth dye, which accumulates in mitochondria proportionally to membrane potential. We observed significantly increased MitoHealth fluorescence in both p.Tyr622* and p.Glu50fs* iNs compared to WT per NeuN surface (Figure~\ref{fig:main_mitochondrial}H). To further confirm these findings in live cells, we measured ΔѰm in soma and neuronal processes using the fluorescent cation tetramethylrhodamine methyl ester (TMRM) in non-quenching mode. TMRM accumulation was higher in p.Tyr622* iNs relative to WT (Figure~\ref{fig:main_mitochondrial}I), and the specificity of this mitochondrial signal was validated through depolarization with the uncoupler FCCP (Figure~\ref{fig:oxygen_consumption_rates_iPSC_neurons}J). Together, these results indicate that ABCA7 LoF iNs exhibit elevated ΔѰm. 

Regulated uncoupling serves as a mechanism to control membrane potential and mitigate oxidative stress \textcolor{red}{[citation needed]}. To assess whether ABCA7 LoF iNs exhibited elevated ROS levels, we incubated p.Tyr622* iNs with Cellrox dye, a fluorescent indicator of oxidative stress. We observed significantly increased Cellrox fluorescence in p.Tyr622* iNs compared to WT iNs (Figure~\ref{fig:main_mitochondrial}J), indicating elevated ROS accumulation in ABCA7 LoF iNs. Together, these data suggest that ABCA7 LoF variants impair mitochondrial uncoupling and increase oxidative stress in neurons.

\subsubsection{ABCA7 LoF induces phosphatidylcholine imbalance in neurons}
\newcommand{\quoteA}{\textcolor{blue}{Since ABCA7 functions as a lipid transporter, we examined the lipidome using LC-MS. Comparing lipidomic profiles between WT and p.Glu50fs3 iNs revealed significant alterations across multiple lipid classes, including neutral lipids, phospholipids, sphingolipids, and steroids (Figure~\ref{fig:main_lipids}A,B). Among these, triglycerides (TGs)—particularly species enriched in long-chain, predominantly polyunsaturated fatty acids—were frequently altered, showing significant upregulation in p.Glu50fs3 iNs (Figure~\ref{fig:main_lipids}B,C).\label{quoteA-label}}}
\quoteA

\newcommand{\quoteB}{\textcolor{blue}{In line with ABCA7's established role as a phospholipid transporter, several phospholipid species exhibited notable differences (Figure~\ref{fig:main_lipids}B). Phosphatidylcholines (PCs), which are essential structural components of biological membranes and potential ABCA7 substrates \textcolor{red}{[citation needed]}, were most prominently affected; 21 out of 32 detected PC species showed increased abundance in p.Glu50fs3 iNs (Figure~\ref{fig:main_lipids}B). Further analysis based on fatty acid saturation—an important factor influencing membrane fluidity—revealed significant enrichment of saturated (SFA) and monounsaturated (MUFA) PCs among the upregulated species (hypergeometric p=0.0027) (Figure~\ref{fig:main_lipids}D). In contrast, polyunsaturated fatty acid-containing (PUFA) PCs were primarily downregulated (p=0.0148) (Figure~\ref{fig:main_lipids}E,F).}}
\quoteB

To determine whether neutral lipid and phosphatidylcholine imbalances were conserved in p.Tyr622* iNs, we performed targeted lipidomic analysis in positive ionization mode. Consistent with p.Glu50fs3 iNs, upregulated lipids in p.Tyr622* iNs were significantly enriched for monounsaturated phosphatidylcholines (hypergeometric p=0.0436) (Figure~\ref{fig:main_mitochondrial}K; Figure~\ref{fig:main_lipids}G,H). Polyunsaturated-containing phosphatidylcholines and long-chain triglycerides were not reliably detected in this LC-MS run (Figure~\ref{fig:main_lipids}I,J), leaving it unclear whether p.Tyr622* iNs exhibit the same extent of polyunsaturated-phosphatidylcholine downregulation or long-chain triglycerides upregulation as p.Glu50fs3 iNs.

De novo phosphatidylcholine synthesis occurs via the Kennedy pathway, followed by subsequent remodeling of the fatty acyl chains by LPCAT enzymes through the Lands cycle, producing phosphatidylcholine species with varying fatty acid saturation levels \cite{Boumann2003-ew;Wang2019-om}. LPCAT3 specifically introduces polyunsaturated fatty acyl chains into PCs \cite{Zhao2008-pq}. We observed decreased LPCAT3 expression in p.Tyr622* and p.Glu50fs*3 iNs compared to WT (Figure~\ref{fig:main_lipids}K,L), aligning with increased levels of saturated PCs in these cells. \newcommand{\quoteH}{\textcolor{blue}{Overall, ABCA7 LoF iNs accumulate neutral lipids, including long-chain polyunsaturated triglycerides and sterol lipids (ZyE), and show imbalances in phosphatidylcholine composition, with higher saturated and monounsaturated species. \label{quoteH-label}}}
\quoteH


\subsubsection{Treatment with CDP-choline reverses impacts of ABCA7 LoF in neurons}
Previous work indicate that exogenous choline supplementation supports the synthesis of unsaturated phosphatidylcholines \cite{Boumann2003-ew}. Consistent with our earlier findings showing that CDP-choline supplementation restores lipid balance and ameliorates lipid-related phenotypes in APOE4 yeast \cite{Sienski2021-zt}, we next examined whether CDP-choline treatment could similarly mitigate metabolic disruptions associated with ABCA7 loss-of-function in iNs.

\newcommand{\quoteD}{\textcolor{blue}{Targeted LC-MS analysis confirmed CDP-choline increased from undetectable to detectable levels in the media following treatment (Figure~\ref{fig:choline_treatment}A). Additionally, both CDP and choline specifically accumulated in media conditioned by p.Tyr622* cells after treatment (Figure~\ref{fig:choline_treatment}A), indicating extracellular hydrolysis of CDP-choline. Intracellular detection of CDP and CDP-choline was not reliable in this experiment. However, intracellular choline levels significantly increased after treatment (Figure~\ref{fig:choline_treatment}B), confirming cellular uptake likely mediated by choline transporters, whose mRNA levels were significantly elevated after treatment (Figure~\ref{fig:main_choline}C), suggesting that choline was successfully taken up by p.Tyr622* iNs upon treatment.\label{quoteD-label}}} 
\quoteD

We anticipated that higher intracellular choline availability would lead to increased phospholipid synthesis. Lipidomic analysis revealed elevated levels of phospholipids, particularly choline-containing lipids (PC, LPC, and SM), alongside reduced levels of neutral lipids (ZyE, TG) (Figure~\ref{fig:main_choline}A). Consistent with these changes, we observed increased expression of PCYT1B, the rate-limiting enzyme in the Kennedy pathway responsible for phosphatidylcholine synthesis (Figure~\ref{fig:choline_treatment}C). Additionally, expression of several LPCAT genes, including LPCAT3, was elevated after treatment (Figure~\ref{fig:choline_treatment}D), coinciding with observed increases in both saturated and unsaturated PC species (Figure~\ref{fig:main_choline}A; Figure~\ref{fig:choline_treatment}D). These findings suggest that CDP-choline treatment promotes synthesis and remodeling of choline-containing lipids.

Transcriptionally, CDP-choline treatment induced significant changes, clearly distinguishing treated from untreated samples (Figure~\ref{fig:choline_treatment}F). The transcriptional signature of CDP-choline treatment negatively correlated with that of p.Tyr622* (Figure~\ref{fig:main_choline}B), suggesting partial restoration toward the WT state. Performing K/L cluster analysis on the p.Tyr622* vs CDP-choline treated p.Tyr622* samples (Figure~\ref{fig:main_choline}C), we observed significant overlap in 6 of the 9 clusters identified in the p.Tyr622* vs WT comparison, with 5 of these clusters showing reversed directional changes following treatment ((Figure~\ref{fig:main_choline}D)). Specifically, clusters related to cell cycle, ribosomal, and proteasomal functions (PM.4, T.12, and T.8)—previously upregulated in p.Tyr622*—were downregulated following CDP-choline treatment (Figure~\ref{fig:main_choline}D). Most notably, the mitochondrial cluster (PM.1), previously exhibiting the strongest overlap with postmortem data, was also reversed following treatment (Figure~\ref{fig:main_choline}D). Further analysis using the MioCarta database confirmed a significant reversal in mitochondrial gene expression, including reduced expression of apoptosis-related genes (BID, CASP3; Figure~\ref{fig:main_choline}F), restoration (upregulation) of the mitochondrial metabolic signature (Table~\ref{tab:choline_mito_genes}), and increased expression of regulators of mitochondrial fusion (MFN2, OPA1), a process which enables high metabolic capacity, dissipation of mitochondrial membrane potential, and mitochondrial biogenesis \textocolor{red}{[citation needed]}. Overall, MioCarta genes were significantly reversed following CDP-choline treatment (Figure~\ref{fig:choline_treatment}G). We observe a significant rescue of the metabolomic signature (Figure~\ref{fig:main_choline}G).

To determine whether CDP-choline treatment could restore mitochondrial uncoupling to WT levels, we repeated the Seahorse assay on p.Tyr622* iNs with and without CDP-choline treatment (Figure~\ref{fig:choline_treatment}H). CDP-choline treatment significantly increased uncoupled respiration in p.Tyr622* iNs, restoring it to WT levels (Figure~\ref{fig:main_choline}J), with no significant change in spare respiratory capacity (Figure~\ref{fig:choline_treatment}J). Consistent with this finding, both TMRM staining (Figure~\ref{fig:main_choline}I) and MitoHealth fluorescence per NeuN-positive surface confirmed that mitochondrial membrane potential (ΔѰm) was significantly reduced in treated cells (Figure~\ref{fig:choline_treatment}K). Additionally, CDP-choline treatment significantly decreased Cellrox fluorescence (Figure~\ref{fig:main_choline}J), indicating a reduction in oxidative stress. 

\subsubsection{CDP-Choline Ameliorates AD-Associated Phenotypes in Cortical Organoids}
Next, we examined whether CDP-choline treatment could mitigate amyloid-beta pathology in a more mature and robust neuronal model. Although ABCA7 LoF iNs showed increased amyloid-beta secretion, absolute concentrations remained very low. Therefore, we differentiated the p.Tyr622* and WT lines into cortical organoids matured for 6 months, a time point selected based on robust amyloid-beta secretion and neuronal maturation, as confirmed by staining for several maturity markers (Figure~\ref{fig:neurospheroid_figure}A). To assess if hyperexcitability—observed previously in iNs—was also present in these organoids, we performed calcium imaging and found significantly increased calcium event frequency in p.Tyr622* organoids (Figure~\ref{fig:neurospheroid_figure}B). Consistent with these findings, cell-attached electrophysiological recordings showed increased spontaneous action potential firing in p.Tyr622* organoids (Figure~\ref{fig:main_choline}K). CDP-choline treatment (1 µM) significantly reversed this electrophysiological hyperexcitability, with a similar trend observed by calcium imaging (Figure~\ref{fig:neurospheroid_figure}B; (Figure~\ref{fig:choline_treatment}K)). Importantly, 1 µM CDP-choline reduced amyloid-beta secretion back to WT levels (Figure~\ref{fig:main_choline}K), an effect that was not observed at lower CDP-choline concentrations or at earlier developmental time points when amyloid-beta secretion was not yet elevated (Figure~\ref{fig:neurospheroid_figure}C).

    