\subsubsection{Extended Data table legends}
\subsubsection{Supplementary Table 1.} 
Annotation of ABCA7 loss-of-function variants identified in this study.\newline
\subsubsection{Supplementary Table 2.} 
Rare ABCA7 loss-of-function variants identified in the ROSMAP cohort (all on chromosome 19).\newline
\subsubsection{Supplementary Table 3.} 
External datasets used. \newline
\subsubsection{Supplementary Table 4.} 
Metadata for ROSMAP cohort individuals analyzed.\newline
\subsubsection{Supplementary Table 5.} 
Technical parameters from Cell Ranger analysis for snRNA-seq samples.\newline
\subsubsection{Supplementary Table 6.} 
Differential gene expression statistics by cell type \newline
\subsubsection{Supplementary Table 7.} 
Summary of gene clusters identified from the UMAP projection of ABCA7 LoF-associated differential expression scores across cell types.\newline
\subsubsection{Supplementary Table 8.} 
Enriched pathways identified by fGSEA (WikiPathways database; cutoff p<0.05) comparing ABCA7 LoF to control excitatory neurons.\newline
\subsubsection{Supplementary Table 9.} 
Differential expression analysis results from mRNA-seq comparing WT, p.Tyr622*+/- CDP-choline-treated, and p.Glu50fs*3 iNs.\newline
\subsubsection{Supplementary Table 10.} 
Enriched pathways identified by fGSEA (WikiPathways database; cutoff adj.p<0.05) comparing WT, p.Tyr622*+/-CDP-choline-treated, and p.Glu50fs*3 iNs.\newline
\subsubsection{Supplementary Table 11.} 
Enriched mitochondrial pathways (MitoCarta) comparing WT vs. p.Tyr622* mRNA.\newline
\subsubsection{Supplementary Table 12.} 
Differential lipid abundance results comparing WT, p.Tyr622*+/- CDP-choline-treated, and p.Glu50fs*3 iNs.\newline
\subsubsection{Supplementary Table 13.} 
Targeted metabolomics of p.Tyr622* iNs and culture media comparing CDP-choline-treated and untreated conditions.\newline
\subsubsection{Supplementary Table 14.} 
Enriched mitochondrial pathways (MitoCarta) comparing p.Tyr622* + H\textsubscript{2}O vs. p.Tyr622* + CDP-choline mRNA.\newline
\subsubsection{Supplementary Table 15.} 
Experimentally determined 3D ABCA7 structures used in molecular dynamics simulations.\newline
\subsubsection{Supplementary Table 16.} 
Primers used for PCR and Sanger sequencing (SS).\newline
\subsubsection{Supplementary Table 17.} 
Antibodies used.\newline

