\subsubsection{Molecular Dynamics Simulations Results} 
The root mean square deviation (RMSD) analysis was performed to assess the conformational stability of ABCA7 in different states and mutations (Figure~\ref{fig:main_excitatory}I; Figure~\ref{fig:md_simulations}A). The RMSD values of the $C_\alpha$ atoms were calculated over the 300 ns simulation period for both the closed and open conformations, with G1527 and A1527 mutations. 

The RMSD plot for ABCA7 in the closed conformation with the G1527 mutation (Figure~\ref{fig:main_excitatory}J, K) exhibited large fluctuations throughout the simulation (Figure~\ref{fig:main_excitatory}L). The distribution of RMSD values was broad, indicating significant conformational variability (Figure~\ref{fig:main_excitatory}L; Figure~\ref{fig:md_simulations}C). The high RMSD values suggest that the G1527 mutation in the closed state induces larger flexibility, leading to substantial deviations from the initial structure. 

In contrast, the RMSD analysis for the closed conformation with the A1527 mutation (Figure~\ref{fig:main_excitatory}J,K) showed minor fluctuations (Fig.2L). The RMSD values were consistently lower than those observed for the G1527 mutation, and the distribution of RMSD values followed a normal distribution (Figure~\ref{fig:main_excitatory}L; Figure~\ref{fig:md_simulations}C). This indicates that the A1527 mutation confers lower local flexibility and greater local stability to the closed conformation of ABCA7. 

For the open conformation of ABCA7, both G1527 and A1527 mutations (Figure~\ref{fig:md_simulations}A,C) exhibited minor RMSD fluctuations (Figure~\ref{fig:md_simulations}D). The RMSD values for these mutations were significantly lower compared to the closed conformation with the G1527 mutation (Figure~\ref{fig:md_simulations}E). The RMSD distributions for both mutations in the open conformation were narrow (Figure~\ref{fig:md_simulations}C; Figure~\ref{fig:md_simulations}D), suggesting local stable conformational behavior. 

Principal component analysis (PCA) was performed to further investigate the conformational dynamics of ABCA7 in the different states and mutations. The PCA of the closed conformation with the G1527 mutation showed a widespread along the first principal component (PC1), reflecting the large conformational space sampled during the simulation (Figure~\ref{fig:main_excitatory}M). This is consistent with the broad RMSD distribution, indicating high flexibility and multiple conformational states. 

In contrast, the closed conformation with the A1527 mutation showed a much more confined spread along PC1 and PC2 (Figure~\ref{fig:main_excitatory}M). This limited spread corresponds to the minor RMSD fluctuations observed, suggesting a locally more stable and less flexible structure. 
For the open conformation, PCA indicated two distinct clusters for both G1527 and A1527 mutations (Figure~\ref{fig:md_simulations}D). These clusters represent two closely related conformational states, indicating that both mutations stabilize the open conformation into two main conformational states with minimal fluctuations. 

\newcommand{\quoteM}{\textcolor{blue}{To further explore the local structural variations induced by a p.Ala1527Gly mutation, we analyzed backbone dihedral angles (phi/psi; $\phi/\psi$) for residues 1517–1537. In the open conformation, Gly1527 consistently occupied the $\alpha$-helical region of the Ramachandran plot throughout the simulation. In contrast, Ala1527 showed two distinct populations within the $\alpha$-helical region (Figure~\ref{fig:md_simulations_2}A), suggesting subtle local conformational differences, yet overall preservation of the $\alpha$-helical structure. These findings align closely with the RMSD analysis, and suggest similar conformational behaviors between the variants in the open state. \label{quoteM-label}}}

\quoteM

\newcommand{\quoteN}{\textcolor{blue}{However, significant structural differences emerged in the closed conformation: Ala1527 displayed two preferred conformations—one within the $\alpha$-helical region and another shifted toward the $\beta$-structure region—while Gly1527 explored a broader range of dihedral angles, indicative of greater structural flexibility (Figure~\ref{fig:md_simulations_2}A,B). This observation corroborates earlier RMSD analyses, emphasizing pronounced structural differences specifically in the closed state, with Gly1527 exhibiting significantly greater conformational flexibility compared to Ala1527.}}

\quoteN

\newcommand{\quoteO}{\textcolor{blue}{To complement backbone angle analysis, we also evaluated secondary structure stability throughout the simulation. In the open state, secondary structure content was comparable between variants, maintaining similar $\alpha$-helical character (Figure~\ref{fig:md_simulations_2}C). Upon transitioning to the closed state, both variants experienced a substantial loss of $\alpha$-helical content across residues 1517–1537. Nonetheless, residues 1520–1525 retained partial $\alpha$-helical structure more robustly in the Ala1527 variant compared to Gly1527 (Figure~\ref{fig:md_simulations_2}C)}}

\quoteO

Finally, structural alignment of the closed-state Gly1527 ABCA7 structure (PDB 8EOP) with the closed-state structures of ABCA1 (PDB 7TBW) and ABCA4 (PDB 7LKZ) reveals that residues corresponding to Gly1527 in ABCA7 (V1646 in ABCA1; I1671 in ABCA4) adopt stable α-helical structures (Figure~\ref{fig:md_simulations_2}E,F). In contrast, the Gly1527 residue in ABCA7 exhibits significant flexibility and lacks defined α-helical structure. Interestingly, our simulations indicate that the Ala1527 variant partially restores this local α-helical conformation in ABCA7 (Figure~\ref{fig:md_simulations_2}C,D). These data suggest that the Gly1527 variant induces local structural changes that differentiate ABCA7 from its close homologs ABCA1 and ABCA4.
