\subsubsection{\underline{Experimental Methods on iNs}} 

\paragraph{rTTA and NGN2 Virus production.}
HEK293T cells (ATCC, Cat\#CRL-3216) were maintained in DMEM/F-12, GlutaMAX (ThermoFisher, Cat\#10565018), 10\% fetal bovine serum (GeminiBio, SKU\#100-106), 1\% MEM Non-essential amino acids (Sigma, Cat\#M7145), 1\% sodium pyruvate (ThermoFisher, Cat\#11360070), and 1\% Penicillin-Streptomycin (GeminiBio, SKU\#400-109). Cells were passaged for maintenance with TrypLE (ThermoFisher, Cat\#12605010) at 70-80\% confluence and reseeded 1:10 in 10 cm tissue culture plates.

For transfection, HEK293T cells were seeded at 5x106 cells per 10 cm plate. Transfection mixtures containing the components required for 3rd generation lentiviral production (per 10 cm dish: 10 µg EF1a-rtTA-Hygro (Addgene \#66810) or pLV-TetO-hNGN2-eGFP-Puro (Addgene \#79823), 5 µg pMDL g/pRRE, 2.5 µg pRSV-Rev, 2.5 µg MD2.G, and 48 µL polyethyleneimine (1 mg/mL) in 600 uL OptiMEM (Fisher, Cat\#51-985-034)). Mixtures were inverted 10X and incubated at RT for 20 min, then added dropwise to the dish. Transfection media was removed 16h later and replaced with 10 mL fresh media. Three days after transfection, media was collected and centrifuged at 3000 xg for 5 min at 4°C to pellet any contaminating cells. Supernatant was transferred to sterile Millex glass ultracentrifuge tubes and centrifuged at 25,000 rpm for 2 hours using a SW32Ti rotor in a Beckman Optima L-90K Ultracentrifuge. The pellets were resuspended in 1 mL PBS per 10 cm plate, and stored at -80°C until use.

\paragraph{Lentivirus-mediated NGN2 induction in iPSCs and drug treatments.}
iPSCs were dissociated into single cell suspension with Cell Dissociation Buffer (Life Technologies, Cat\#13151-014), centrifuged at 300 xg for 5 min, and resuspended in mTeSR1 media with Rock inhibitor (Rockout; Abcam, ab285418). Single-cell suspension was plated in a 6-well plate coated with Matrigel for an optimized seeding density of 50-60\% confluence 24 hours after plating. One day after plating, cells were co-transduced with 80 µL pLV-TetO-hNGN2-eGFP-Puro and 80 µL EF1a-rtTA-Hygro added in 1 mL fresh media per well and incubated overnight at 37°C. NGN2 expression was then induced 24 hours later with addition of 2 mL fresh media supplemented with doxycycline (DOX, 1 µg/mL, final concentration) and Rock inhibitor. Puromycin (1 µg/mL) selection of non-NGN2 expressing cells was performed with media change 24 hours after induction, with continued DOX supplementation. After 24 hours of puromycin selection, immature neurons were re-plated onto PDL/laminin coated plates at 1x106 cells/well on 6-well plates or 5x104 cells/well on 96-well plates. Neurons were maintained in BrainPhys Neuronal Media (STEMCELL Technologies, Cat\#05793) with Neurocult SM1 Neuronal Supplement (STEMCELL Technologies, Cat\#05711), N2-supplement-A (STEMCELL Technologies, Cat\#07152), laminin (1ug/mL), and DOX (1 µg/mL) with half media changes every 3-4 days. Neuronal cultures were maintained for 28 days before experimentation.

iPSC-derived neurons were treated with cytidine 5’-diphosphocholine (CDP-choline; Millipore Sigma 30290) to final concentrations of 100 µM beginning at day 14 and repeated with each media change until 28 days matured. Choice of treatment concentration and duration was based on a previous study by our lab \supercite{Sienski2021-zt}.

\paragraph{Electrophysiological recordings.}
Cells were placed in a recording chamber and perfused with oxygenated artificial cerebrospinal fluid (ACSF) contains (in mM) 125 NaCl, 2.5 KCl, 1.2 NaH2PO4•H2O, 2.4 CaCl2•2H2O, 1.2 MgCl2•6H2O, 26 NaHCO3 and 11 D-Glucose at a constant rate of 2 mL/min at ~32°C. Cells were visualized using infrared differential interference contrast (IR-DIC) imaging on an Olympus BX-50WI microscope.

Recordings were performed using Axon Multiclamp 700B and Clampex 11.2 (Molecular Devices). Action potentials were generated by injecting various steps of currents using current clamp configuration.  Whole-cell currents were recorded from a holding potential of -80 mV by stepping to various voltages using voltage clamp configuration. Cell-attached recording configution was used to measure spontaneous firing activity. Signals were filtered at 1 kHz using the amplifier’s four-pole, low-pass Bessel filter, digitized at 10 kHz with a Digidata 1550B interface (Molecular Devices). Pipette solution contained (in mM) 120 K gluconate, 5 KCl, 2 MgCl2•6H2O, 10 HEPES, 4 ATP, 0.2 GTP. pClamp 11.2 (Molecular Devices) and GraphPad Prism 10 software suites were used for data acquisition and analysis. Data are presented as means ± standard errors of means (SEM).

\paragraph{A$\beta$ Enzyme linked immunosorbent assays on iNs.}
Media was collected from 4 week old iNs and flash frozen. ELISAs were performed on thawed media according to manufacturer’s instructions to measure Aꞵ40 (ThermoFisher Scientific, KHB3481) and Aꞵ42 (ThermoFisher Scientific, KHB3441) respectively.

\paragraph{Mitochondrial Health cell stain.}
HCS Mitochondrial Health kit (ThermoFisher, Cat\#H10295) were used on live cells according to manufacturer’s protocols. In brief, 50 uL of media containing 1.5 µL MitoHealth dye was added to each well and incubated on live cells for 30 min at 37°C. Next, neurons were fixed in 4\% paraformaldehyde/4\% sucrose in PBS at 4°C for 15 min at room temperature, washed 3X with PBS, then permeabilized with 0.1\% Triton-X in PBS for 5 min at room temperature. Cells were blocked in 2\% Bovine Serum Albumin (BSA, Fisher Bioreagents, BP9703) in PBS for 20 min at room temperature, then incubated in primary antibodies diluted in blocking solution (NeuN; 1:500) overnight at 4°C. Cells were washed 3X for 5 min with PBS, then incubated in secondary antibodies diluted in blocking solution (1:1000) for 2 hours at room temperature. Cells were washed 3X for 5 min with PBS, then incubated for 10 min with 1:2000 Hoechst 33342 (Invitrogen, H3570). Cells were washed 1X with PBS, and wells flooded with PBS for imaging.\\
Confocal images were acquired on a Zeiss LSM900. Acquisition settings were kept constant within each imaging batch (where conditions of interest were uniformly distributed across plates). The minimum and maximum z-plane was manually determined for each culture well, to accommodate differences in culture thickness. Cultures were imaged at 1 µm intervals along the z-axis.

\paragraph{Live imaging of TMRM.}
Stock solutions were initially prepared at concentrations of 10 µM TMRM (ThermoFisher, I34361), 100 µM FCCP (Cayman Chemical, 15218), and 100 µM oligomycin from Streptomyces diastatochromogenes (Sigma, O4876-25MG), achieving final working concentrations of 0.1 µM TMRM, 1 µM FCCP, and 1 µM oligomycin. For imaging, 3 µL of the 10 µM TMRM stock solution was added to each well of iNs cultured in 300 µL of media on a 96-well plate. Cells were incubated with TMRM dye for 30 minutes at 37°C. Live-cell imaging was then performed using a Zeiss LSM 900 confocal microscope equipped with ZEN software. After initial TMRM images were acquired, 3 µL of the 100 µM FCCP stock was added to each well, and images were captured immediately. Subsequently, 3 µL of the 100 µM oligomycin stock was added, followed by immediate image acquisition. Confocal images were acquired as single optical sections (single-plane imaging) on a Zeiss LSM900 microscope.

\paragraph{Live imaging of CellROX.}
iNs were maintained in 300 µL of media per well in a 96-well plate. A stock solution of CellROX Orange Reagent (ThermoFisher, C10443) was prepared at a concentration of 500 µM. For staining, 3 µL of the 500 µM CellROX dye was added directly to each well containing iNs in 300 µL media. Cells were incubated with the dye for 30 minutes at 37°C. After incubation, live-cell images were acquired as single optical sections (single-plane imaging) on a Zeiss LSM900 microscope.

\paragraph{Seahorse Metabolic Assays.}
iPSCs were differentiated as described above directly on 96-well Agilent Seahorse XFe96/XF Pro cell culture microplates and matured for 28 days before assaying on a Seahorse XFe96 Analyzer. Seahorse XF Cell Mito Stress Test and Oxidation Stress Tests were performed according to manufacturer protocol with the following final drug concentration: Oligomycin, 2.5 µM; FCCP, 1 µM; Rotenone/Antimycin, 0.5 µM. 

\paragraph{MAP2 and NeuN staining of iNs.}
iNs were plated on coverslips with coating. Cells were fixed in 4\% formaldehyde for 10 min at room temperature followed by PBS wash once. Cells were next permeabilized and blocked in PBS containing 0.2\% TritonX-100 and 10\% bovine serum albumin for 1 hour at room temperature, then incubated with primary antibody (MAP2 1:1000; NeuN 1:1000; diluted in blocking solution) at 4 °C overnight. Primary antibody was visualized using the appropriate secondary antibody conjugated to Alexa Fluor 594, or Alexa Fluor 647 (1:500, Thermo Fisher Scientific). Nuclei were visualized with Hoechst 33342 (1:1000, Thermo Fisher Scientific). Coverslips were then mounted on the glass slides with fluoromount g. Confocal images were acquired as single optical sections (single-plane imaging) on a Zeiss LSM900 microscope.

\paragraph{RNA Extraction, Library Preparation, and Sequencing.}
Total RNA was extracted from iNs using the RNeasy Mini Kit (Qiagen), and RNA quality was assessed using a 5300 Fragment Analyzer (Agilent). Only samples with an RNA Quality Number (RQN) greater than 9.5 were selected for library preparation. Full-length cDNA was synthesized using the SMART-seq v4 kit (Takara Bio), and sequencing-ready libraries were subsequently prepared using the Nextera XT DNA Library Preparation Kit (Illumina). Libraries were sequenced as 75 bp + 75 bp paired-end reads with dual 8-nucleotide indexes on an Element AVITI sequencing platform (Element Biosciences) at the MIT BioMicro Center. 